---
title: "Zindi_Traffic Jam analysis"
author: "Denis"
date: " `r as.Date(Sys.time())` "
output:
  pdf_document:
    fig_caption: yes
    number_sections: yes
    toc_depth: 3
  html_document:
    css: kable.css
    toc: yes
---

\newpage
\centering
\raggedright
\newpage
\hypersetup{linkcolor=black}
\tableofcontents
\newpage



```{r echo=F}
library(dplyr)
library(ggplot2)
library(tidyverse)
library(lubridate)
require(randomForest)
library(caret)
```



```{r }

# Reading the data

tr <- read.csv("C:/Users/dmunene/OneDrive - Dalberg Global Development Advisors/RESSOL/Personal/Data Analysis/Traffic Jam/train_revisedV2.csv")
tt <- read.csv("C:/Users/dmunene/OneDrive - Dalberg Global Development Advisors/RESSOL/Personal/Data Analysis/Traffic Jam/test_questions.csv")

tr <- as_factor(tr)
tt <- as_factor(tt)

# Previewing the top data points

head(tr[,])
head(tt[,])


```


#### Processing Date/Time

```{r, echo=FALSE}
# travel date to day of week


tr$travel_date <- lubridate::dmy(tr$travel_date)
tr$travel_date<-  weekdays(tr$travel_date)

tt$travel_date <- lubridate::ymd(tt$travel_date)
tt$travel_date <-  weekdays(tt$travel_date)

# Parsing time and creating time in Hour
tr$travel_time <- parse_time(tr$travel_time)
tr$travel_timeH <- lubridate::hour(tr$travel_time)

tt$travel_time <- parse_time(tt$travel_time)
tt$travel_timeH <- lubridate::hour(tt$travel_time)
```





```{r, echo=F}
# Grouping data on number of tickets and merging with main dataset

tr1 <- tr %>% group_by(ride_id,travel_date) %>% summarise(number_of_tickets=n())

tr2 <- merge(tr1,tr)



tt1 <- tt %>% group_by(ride_id,travel_date) %>% summarise(number_of_tickets=n())

tt2 <- merge(tt1,tt)




# selecting variables for analysis

tr3 <- tr2 %>% dplyr:: select(travel_date,travel_timeH,travel_from,car_type,max_capacity,number_of_tickets)

tt3 <- tt2 %>% dplyr:: select(ride_id,travel_date,travel_timeH,travel_from,car_type,max_capacity,number_of_tickets)

anyNA(tr3)

```



```{r echo=T}

#divide the data into train and test
set.seed(121)

trainRowNumbers <- createDataPartition(tr3$number_of_tickets, p = 0.7, list = F)


train_dt <- tr3[trainRowNumbers,] 
test_dt <- tr3[-trainRowNumbers,]



# Store X and Y for later use.
x <- train_dt[,1:5]
y <- train_dt$number_of_tickets

# One-Hot Encoding
# Creating dummy variables is converting a categorical variable to as many binary variables as here are categories.

dummies_model <- dummyVars(number_of_tickets~., data = train_dt)

# Create the dummy variables using predict. The Y variable (number_of_tickets) will not be present in train_dt2.
train_dt <- predict(dummies_model,newdata = train_dt)

# # Convert to dataframe
train_dt <- data.frame(train_dt)

# # See the structure of the new dataset
str(train_dt)


# Controliing for the range to between 0 and 1
preProcess_range_model <- preProcess(train_dt, method = 'range')
train_dt <- predict(preProcess_range_model, newdata = train_dt)

# Append the Y variable
train_dt$number_of_tickets <- y

apply(train_dt[,1:10],2,FUN = function(x){c('min' = min(x),'max' = max(x))})

```



### Train in model mars

```{r}

# Set the seed for reproducibility
set.seed(134)

# Train the model using randomForest and predict on the training data itself.
model_mars = train(number_of_tickets ~., data = train_dt, method = 'earth')
fitted <- predict(model_mars)

model_mars

plot(model_mars, main = "Model Accuracies with MARS")

varimp_mars <- varImp(model_mars)
plot(varimp_mars, main = "Variable Importance with MARS")


```


```{r}

# Create one-hot encodings (dummy variables) on test data
tr_test <- predict(dummies_model,test_dt)

tr_test <- data.frame(tr_test)

preProcess_range_model2 <- preProcess(tr_test, method = 'range')

tr_test <- predict(preProcess_range_model2,newdata =tr_test)

head(tr_test[,1:7])


```



```{r}


# Predict on testData

predicts <- predict(model_mars,tr_test)

prr <- round(predicts)

head(prr)


```

### Testing the MAE

```{r}
MAE(prr,test_dt$number_of_tickets)

```

## Preparing test data

```{r echo=T}
# Step 2: Create one-hot encodings (dummy variables) on test data
tr_test <- predict(dummies_model,test_dt)

tr_test <- data.frame(tr_test)

preProcess_range_model2 <- preProcess(tr_test, method = 'range')

tr_test <- predict(preProcess_range_model2,newdata =tr_test)

head(tr_test[,1:7])


```



```{r}

# Step 2: Create one-hot encodings (dummy variables) on MAIN test data
dummies_model2 <- dummyVars(number_of_tickets~., data = tt3)

tt_dt <- predict(dummies_model2,newdata = tt3)


tt_dt <- data.frame(tt_dt)


preProcess_range_model3 <- preProcess(tt3, method = 'range')



# Step 3: Transform the features to range between 0 and 1
tt4 <- predict(preProcess_range_model3, tt_dt)

# View
head(tt4[,1:7])

```





### Train in SVM

```{r}

model_svm <- train(number_of_tickets~., data = tr3, method = 'svmRadial')

model_svm

```





### Train in random forest

```{r}

# Create a Random Forest model with default parameters

model1 <- randomForest(number_of_tickets ~ ., data = train_dt, importance = T)

model1



```



```{r}

# Predict on testData

predictss <- predict(model1,tr_test)

prrr <- round(predictss)

head(prrr)


model3 <- randomForest(number_of_tickets ~ ., data = train_dt, ntree = 500, mtry = 12, importance = T)

model3


# Predict on testData

predict_frst3 <- predict(model3,tr_test)

p_forest3 <- round(predict_frst3)

head(p_forest3)

```




### Testing the MAE for random forest

```{r}
MAE(prrr,test_dt$number_of_tickets)

MAE(p_forest3,test_dt$number_of_tickets)

```

### Running model on main test set
```{r}

prediction2 <- predict(model1,tt4)

p_forest <- round(prediction2)

head(p_forest)


```


### Running model3 on main test set

```{r}

predict_frst3 <- predict(model3,tt4)

p_forest3 <- round(predict_frst3)

head(p_forest3)


```



## Embedd prediction to main test data and output submission

```{r}
# convert to dataframe
p_forest <- data.frame(p_forest)


# embedding on main test data


df <- data.frame(tt2$ride_id,p_forest)
df_f <- df %>%
  select(ride_id = tt2.ride_id,number_of_tickets=p_forest)

write.csv(df_f,"C:/Users/dmunene/OneDrive - Dalberg Global Development Advisors/RESSOL/Personal/Data Analysis/Traffic Jam/Nesh_Submission3.csv", row.names = F)



```


















